Realizing Temporal Graphs

### 2023-01-13 RealizingTemporalGraphs.tex
Final version _ 02 :)

\subsection{Step 1 of the FPT algorithm}
In the first step of the algorithm we create an instance $D'$, with the underlying graph $G'$, from $D$ and $G$, by simply removing all the vertices of degree one from $G$ and their corresponding rows and columns from the matrix $D$. 
This part is executed in $O(n+m)$ time.

@@

Let $s \in V(G) \setminus V(G')$ be a vertex of degree one that was removed from $G$ at the beginning,
and let $u$ be its neighbor in $G$.
First, we can easily assume that $|N_{G'}(u)|\geq 1$. If this is not true then either $G$ is a star or $G'$ is disconnected and therefore also $G$ is disconnected, where $u$ is in a part of $G$ that forms a star.iteratively
We can solve the problem for stars in polynomial time (see \cref{thm:deltaExact-PolyTimeTrees}).

To determine the correct label of edge $su$ we use the following properties.
\begin{claim}
Let $v_i \in N_{G'}(u)$ be the closest vertex to $v$, regarding the values of $D$, \ie  $D_{s,v_i} \leq D_{s,v_j}$ for all $v_j \in N_{G'}(u)$.
Then the path $P=(s,u,v_i)$ has to be the fastest temporal path from $s$ to $v_i$ in $G$.
\end{claim}
\begin{proof}
Suppose that this is not true.
Then there exists a faster path $P^*$ from $s$ to $v_i$, that goes through $u$ (as this is the only neighbor of $s$), through some other vertex $v_j \in N_{G'}(u) \setminus \{v_i \}$, 
and through some other path $P'$ in $G$, before it finishes in $v_i$, where $P'$ is at least an edge (from $v_j$ to $v_i$). 
Therefore $P^* = (s,u,v_j,P',v_i)$, where $d(P^*) \leq d(P)$.
Now since $D_{s,v_i} \leq D_{s,v_j}$ for all $v_j \in N_{G'}(u)$ the first part of path $P^*$ from $s$ to $v_j$ takes at least $D_{s,v_i}$ time.
Since $v_i \neq v_j$ we need at least one more time-step (one more edge) to traverse from $v_i$ to reach $v_j$. 
So $d(P^) \geq D_{s,v_i} + 1$, and so $P^*$ cannot be faster than $P$.
\end{proof}

Since $P=(s,u,v_i)$ is the fastest temporal path from $s$ to $v_i$ we calculate the label of edge $su$ as $\lambda (s,v_i) \equiv \lambda(u v_i) + 1 - D_{s,v_i} \pmod \Delta$.
We repeat the above procedure for all degree one vertices in $G$, that we removed in Step 1, and we end up with the labeling $\lambda$ of the whole graph.

>>>>

\subsection{Step 1 of the FPT algorithm}
In the first step of the algorithm we create an instance $D'$, with the underlying graph $G'$, from $D$ and $G$, by iteratively removing vertices of degree one from $G$ and their corresponding rows and columns from the matrix $D$. 
This part is executed in $O(n+m)$ time.

@@

Note that vertices from $V(G) \setminus V(G')$ form disjoint paths in $G$. 
Each such path is incident to exactly one vertex from $G'$.

Let $P' = (v_1, v_2, \dots, v_p=s)$ be a path in $G$ on vertices from $V(G) \setminus V(G')$  (\ie path on vertices that were removed from $G$)
and let $u$ be a neighbor of $s$ in $G$,
therefore $P=(v_1, v_2, \dots, v_p=s, u)$ is an induced path in $G$.
First observe that $|N_{G'}(u)|\geq 1$, if not also $u$ would be removed from $G$ and $G$ would be a tree, for which we can solve the problem in polynomial time (see~\cref{thm:deltaExact-PolyTimeTrees}).

To determine the correct label of all edges of $P$ we use the following property.
\begin{claim}
Let $P = (v_1, v_2, \dots, v_p=s)$ be a path in $G$ on vertices from $V(G) \setminus V(G')$
and let $u$ be a neighbor of $s$ in $G$.
Let $v_i \in N_{G'}(u)$ be the closest vertex to $v$, regarding the values of $D$, \ie  $D_{s,v_i} \leq D_{s,v_j}$ for all $v_j \in N_{G'}(u)$.
Then the path $P^*=(s,u,v_i)$ has to be the fastest temporal path from $s$ to $v_i$ in $G$.
\end{claim}
\begin{proof}
Suppose that this is not true.
Then there exists a faster path $P^*_2$ from $s$ to $v_i$, that goes through $u$ (as this is the only neighbor of $s$), through some other vertex $v_j \in N_{G'}(u) \setminus \{v_i \}$, 
and through some other path $P'$ in $G$, before it finishes in $v_i$, where $P'$ is at least an edge (from $v_j$ to $v_i$). 
Therefore $P^*_2= (s,u,v_j,P',v_i)$, where $d(P^*_2) \leq d(P^*)$.
Now since $D_{s,v_i} \leq D_{s,v_j}$ for all $v_j \in N_{G'}(u)$ the first part of path $P^*_2$ from $s$ to $v_j$ takes at least $D_{s,v_i}$ time.
Since $v_i \neq v_j$ we need at least one more time-step (one more edge) to traverse from $v_i$ to reach $v_j$. 
So $d(P^*) \geq D_{s,v_i} + 1$, and so $P^*_2$ cannot be faster than $P^*$.
\end{proof}

Because $(s,u,v_i)$ is the fastest temporal path from $s$ to $v_i$ we calculate the label of edge $su$ as $\lambda (s,v_i) \equiv \lambda(u v_i) + 1 - D_{s,v_i} \pmod \Delta$.
Since $P$ is a path, the fastest temporal path from all vertices of $P$ to $u$ is the temporal path along $P$, therefore using \cref{obs:FirstLabelAndDuration} and the value of the edge $su$, we can determine labels of all edges in $P$.

We repeat the above procedure for paths in $G$ that we removed in Step 1, and we end up with the labeling $\lambda$ of the whole graph.
